\chapter{Place and Route}

For what concerns the place \& route, we have followed the guidelines provided for the first laboratory activity, hence there is nothing particular to be explained.

\section{Timing report}

Here it is reported the reg-to-reg analysis provided by Innovus after the RISC-V basic design has been placed and routed.

\begin{Verbatim}[fontsize=\small]
Path 1: MET Setup Check with Pin d/ids/reg_file/reg_i_24/q_reg[1]/CK 
Endpoint:   d/ids/reg_file/reg_i_24/q_reg[1]/D (^) checked with trailing edge 
of 'my_clk'
Beginpoint: c/curr_wb_reg[1]/Q                 (^) triggered by  leading edge 
of 'my_clk'
Path Groups: {reg2reg}
Analysis View: MyAnView
Other End Arrival Time          1.350
- Setup                         0.042
+ Phase Shift                   0.000
- Uncertainty                   0.070
= Required Time                 1.238
- Arrival Time                  1.049
= Slack Time                    0.189
Clock Rise Edge                      0.000
= Beginpoint Arrival Time            0.000
Timing Path:
+-------------------------------------------------------------------------------------------+ 
|                Pin                 | Edge |  Net  |   Cell   | Delay | Arrival | Required | 
|                                    |      |       |          |       |  Time   |   Time   | 
|------------------------------------+------+-------+----------+-------+---------+----------| 
| clk                                |  ^   | clk   |          |       |   0.000 |    0.189 | 
| c/curr_wb_reg[1]/CK                |  ^   | clk   | DFF_X1   | 0.000 |   0.000 |    0.189 | 
| c/curr_wb_reg[1]/Q                 |  ^   | wb_en | DFF_X1   | 0.098 |   0.098 |    0.287 | 
| U1945/A3                           |  ^   | wb_en | NAND3_X1 | 0.000 |   0.098 |    0.287 | 
| U1945/ZN                           |  v   | n1509 | NAND3_X1 | 0.060 |   0.157 |    0.346 | 
| U2625/B1                           |  v   | n1509 | OAI21_X1 | 0.001 |   0.158 |    0.347 | 
| U2625/ZN                           |  ^   | n1529 | OAI21_X1 | 0.493 |   0.651 |    0.840 | 
| U2349/A2                           |  ^   | n1529 | NAND2_X1 | 0.006 |   0.658 |    0.847 | 
| U2349/ZN                           |  v   | n1530 | NAND2_X1 | 0.277 |   0.935 |    1.124 | 
| U3821/B2                           |  v   | n1530 | OAI22_X1 | 0.027 |   0.962 |    1.151 | 
| U3821/ZN                           |  ^   | n2053 | OAI22_X1 | 0.087 |   1.049 |    1.238 | 
| d/ids/reg_file/reg_i_24/q_reg[1]/D |  ^   | n2053 | DFF_X1   | 0.000 |   1.049 |    1.238 | 
+-------------------------------------------------------------------------------------------+ 
Clock Fall Edge                      1.350
= Beginpoint Arrival Time            1.350
Other End Path:
+----------------------------------------------------------------------------------------+ 
|                 Pin                 | Edge | Net |  Cell  | Delay | Arrival | Required | 
|                                     |      |     |        |       |  Time   |   Time   | 
|-------------------------------------+------+-----+--------+-------+---------+----------| 
| clk                                 |  v   | clk |        |       |   1.350 |    1.161 | 
| U1796/A                             |  v   | clk | INV_X2 | 0.000 |   1.350 |    1.161 | 
| U1796/ZN                            |  ^   | n59 | INV_X2 | 0.000 |   1.350 |    1.161 | 
| d/ids/reg_file/reg_i_24/q_reg[1]/CK |  ^   | n59 | DFF_X1 | 0.000 |   1.350 |    1.161 | 
+----------------------------------------------------------------------------------------+ 
\end{Verbatim}

As it can be seen, the critical path is different from the one reported by the synthesis tool. In fact,
the critical path obtained with the synthesis was in the EXE adder, while here it starts from the WB stage
and arrives to the register file. We think that this difference is due to the fact that Innovus takes into
consideration the delay of the nets, since it knows exactly their length and where their position.
Moreover, this critical path makes sense because the logic is constrained to work in half clock cycle, because
the register file samples the data on the clock's falling edge.

The critical path for the enhanced version of the processor is similar, as it arrives to the same register
but to a different flip-flop, hence it will not be reported here.

\section{Area report}

The area reported by Innovus is a bit smaller than the one reported by Synopsys. For the basic RISC-V processor
we obtained an area occupation of $11817.8\ \mu m^2$, while for the processor with the \verb|ABS| unit the 
area is $12172.7\ \mu m^2$. This reduction is expected, since after the place and route flow the size is no more
an estimation and the tool is able to better optimise the placing of the cells.